### 2022-12-05 intro.Rnw
change paper order 1 and 2
table formatting

\hfill
\begin{tabular}{p{0.6\textwidth}}
  ``\emph{Confirmation comes from repetition}. Any attempt to avoid this statement leads at least to failure and more probably to destruction.'' \\
  \hfill \citet[p. 84]{Tukey1969}
\end{tabular}

@@

% How can we know if a finding from a scientific study is really true? For
% example, how can we know if the protective effect of a vaccine found in a study
% is real? The answer to such questions is of high importance to scientists,
% decision makers, and the general population. However, because any study result
% comes with uncertainty we can never know for sure.
How can we know whether we can trust a finding from a scientific study? For
example, how can we know whether the protective effect of a vaccine found in a
study is a genuine effect or just a spurious finding? The answer to such
questions is of high importance to scientists, decision makers, and the general
population. However, because any study result comes with uncertainty we can
never know for sure.

One way to come a little closer to the truth is to repeat the original study
with new subjects. Such a \emph{replication study} may yield similar results
which would make us more confident about the original finding, or may yield
conflicting results which would lower our confidence. Replication studies are
thus an essential part of the scientific process as they provide a means for
substantiating genuine research findings and refuting research findings which
occurred merely by chance. For this reason, ``successful'' replication is often
a requirement, for example, for acceptance of newly proposed scientific theories
(\eg{} a new physical model) or the implementation of policies based on
scientific knowledge (\eg{} market approval of a drug). The results from
replication studies may thus have real world consequences, such as deciding
whether we receive the recommendation to get vaccinated (a consequence that
anyone who experienced the COVID-19 pandemic is very aware of).

Despite their tremendous importance, the traditional academic system has made it
unattractive for researchers to conduct replication studies; the currencies of
science -- publications, citations, and grant money -- are typically easier to
acquire by conducting novel research. This is because until relatively recently,
many scientific journals put most emphasis on ``innovation'' thereby lowering
the chance of publication for replication studies \citep{Makel2012, Martin2017,
  Coiera2021}. To build a successful career, researchers thus often had no other
choice than to concentrate their efforts on producing novel and eye-catching
research results.

@@

Most of these large-scale replication projects confirmed what many researchers
had feared; carefully conducted replication studies often show less impressive
results than their original counterparts, and the replicability of research
findings is surprisingly low (whatever replicability means will soon be
discussed in more detail). This realization led many to declare science as being
in a ``replication crisis''. Debates arose about whether or not the crisis
really existed and who or what was to blame \citep{Gilbert2016, Amrhein2019}.
Even the popular press became interested \citep[\eg{}][]{Carey2015, Kovic2016,
  Achenbach2017, Devlin2018}, so that also in the eyes of the public the
credibility of science became seriously threatened.

\begin{figure}[!htb]
<< "google-scholar-graph", fig.height = 2.5 >>=
## Data gathered on October 18 2022
searchTerm <- "Replication Study"
searchDate <- as.Date(x = "2022-10-18", format = "%Y-%m-%d", tz = "GMT")
searchDateFormat <- format(searchDate, format = "%B %d %Y")
resultsAnytime <- 114000
scholarDF <- data.frame(year = seq(2022, 1990, -1),
                        results = c(4530, 11400, 13600, 12200, 11300, 10700,
                                    9880, 9220, 8660, 8200, 7500, 6700, 6040,
                                    5370, 4680, 3890, 3380, 3070, 2790, 2520,
                                    2310, 2190, 2120, 2020, 1920, 1840, 1730,
                                    1490, 1270, 1230, 1140, 1070, 970),
                        study = c(156000, 487000, 1820000, 959000, 2360000,
                                  675000, 3000000, 688000, 3150000, 512000,
                                  3630000, 451000, 3650000, 396000, 3680000,
                                  389000, 3890000, 375000, 3950000, 336000,
                                  4160000, 364000, 4240000, 385000, 3470000,
                                  377000, 3120000, 335000, 2620000, 376000,
                                  2440000, 317000, 2220000))

## plot data
ggplot(data = filter(scholarDF, year < 2022),
       aes(x = year, y = results)) +
    geom_segment(aes(xend = year, yend = 0), alpha = 0.2, size = 0.4) +
    geom_point() +
    ## scale_y_continuous(breaks = seq(0, 15000, 2500)) +
    scale_x_continuous(breaks = seq(1990, 2020, 5)) +
    ## scale_y_log10() +
    labs(x = "Year", y = paste0('Number of results \nfor "', searchTerm, '"')) +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank()## ,
          ## panel.grid.major.y = element_line(colour = "grey", linetype = "dashed")
          )
@
\caption{Number of results per year for search term ``\Sexpr{searchTerm}'' on
  Google Scholar. The search was conducted on 18 October 2022.}
\label{fig:scholar}
\end{figure}

In the midst of this crisis, % , particularly from the hard-hit
% field of psychology,
various reforms were implemented to prevent an escalation of the situation % and
% save the reputation of science
\citep[for an overview see \eg{}][]{Munafo2017}. For instance, many journals
adopted the ``registered report'' format \citep{Chambers2021} in which a study
proposal is peer reviewed \emph{before} the study is conducted, and which, if
accepted, gives provisional publication acceptance regardless of the study
outcome. Similarly, digital infrastructure platforms, such as Zenodo
(\url{https://zenodo.org}) or the Open Science Framework
(\url{https://osf.io}), were created to facilitate preregistration, preprints,
code and data sharing, all of which have substantially increased over the last
decade \citep{Kidwell2016, Nosek2018, Rawlinson2019}. The practice of conducting
replication studies has also gained popularity. Several academic journals and
funders are now explicitly promoting replication research \citep{NWO2016,
  NSL2018, Nature2022}. Figure~\ref{fig:scholar} illustrates this trend via the
yearly number of results for the search term ``\Sexpr{searchTerm}'' on Google
Scholar over the last three decades. We see that the numbers have been rapidly
growing, especially after the mid 2000s, with a minor drop in 2021 (perhaps
because of research slowing down due to the COVID-19 pandemic).

@@

\begin{table}[!htb]
  \centering
  \caption{Study-level summary statistics for an original and a replication
    study. The cumulative distribution function of the standard normal
    distribution is denoted by $\Phi(\cdot)$, and the $1 - \alpha$ quantile of
    the standard normal distribution is denoted by
    $\Phi^{-1}(1 - \alpha) = \z{\alpha}$. Confidence intervals and
    \textit{p}-values are based on the normality assumption.}
  \label{tab:meta}
  \begin{tabular}{lcc}
    \toprule
    & Original study & Replication study \\
    \midrule
    effect estimate & $\that_o$ & $\that_{r}$ \\
    standard error & $\sigma_{o}$ & $\sigma_{r}$ \\
    $(1 - \alpha)$ confidence interval &
    $[\that_{o} \pm \z{\alpha/2} \sigma_{o}]$ &
    $[\that_{r} \pm \z{\alpha/2} \sigma_{r}]$ \\
    % $[\that_{o} - \z{\alpha/2} \sigma_{o}, \that_{o} + \z{\alpha/2} \sigma_{o}]$ &
    % $[\that_{r} - \z{\alpha/2} \sigma_{r}, \that_{r} + \z{\alpha/2} \sigma_{r}]$ \\
    $z$-value & $z_{o} = \that_{o}/\sigma_{o}$ & $z_{r} = \that_{r}/\sigma_{r}$ \\
    $p$-value (two-sided) & $p_{o} = 2\{1 - \Phi(|z_{o}|)\}$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    % One-sided $p$-value & $p_{o} = 1 - \Phi(|z_{o}|)$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    \bottomrule
  \end{tabular}
\end{table}

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in research articles, and if one of them is
missing they typically can be calculated from the others. Using summary
statistics is also often the only possible way for conducting the replication
success analysis as the raw data from the original study may not be available to
the researchers conducting the replication study. The most important statistic
is the effect estimate $\that$. It provides an estimate of the underlying effect
size $\theta$ which quantifies the true effect or association of an
intervention/exposure with an outcome variable. Typical effect sizes are mean
differences and correlations (for continuous outcomes), odds ratios (for binary
outcomes), or hazard ratios (for time to event outcomes). Depending on the type
of effect size, a transformation might be required so that the assumption of
approximately normally distributed effect estimates around the unknown effect
size $\theta$ (for large enough sample sizes) is justified. This could be, for
instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard error $\sigma$ represents the statistical uncertainty of the
estimate. Under the assumption of (asymptotic) normality, confidence intervals
for $\theta$ and $p$-values for testing the null hypothesis
$H_{0} \colon \theta = 0$ can be computed as shown in Table~\ref{tab:meta}.

\begin{table}[!htb]
  \centering
  \caption{Statistical criteria for assessing replication success which have
    been used in practice \citep{Opensc2015, Camerer2016, Camerer2018, Cova2018,
      Errington2021}.}
  \label{tab:rsdef}
  \begin{tabular}{p{0.21\textwidth}p{0.7\textwidth}}
    \toprule
    Criterion type & Description \\
    \midrule
    Significance & The original and replication $p$-values are smaller than
                   a threshold $\alpha$ and their effect estimates show
                   the same direction
                   (\mbox{$p_{o} < \alpha$}, \mbox{$p_{r} < \alpha$}, and
                   \mbox{$\sign(\that_{o}) = \sign(\that_{r})$}). Usually $\alpha = 5\%$.
    \\
                   & \\
    Meta-analytic significance & The meta-analytic $p$-value is smaller than a threshold
                                 $\alpha$ (\mbox{$p_{m} = 2\{1 - \Phi(|\that_{m}|/\sigma_{m})\} < \alpha$})
                                 where \mbox{$\that_{m} =
                                 (\that_{o}/\sigma_{o}^{2} + \that_{r}/\sigma^{2}_{r})\sigma^{2}_{m}$}
                                 is the  pooled effect estimate with standard error
                                 \mbox{$\sigma_{m} = (1/\sigma^{2}_{o} + 1/\sigma^{2}_{r})^{-1/2}$}.
                                 Usually $\alpha = 5\%$. \\
                   & \\
    Relative effect size & The effect estimate of the replication study goes in the same
                           direction as the original one and is at least $d_{\text{min}}$ times as large (\mbox{$d = \that_{r}/\that_{o} \geq d_{\text{min}}$}).
                           Usually $d_{\text{min}} = 1$. \\
                   & \\
    Confidence interval & The replication effect estimate is contained in the
                          $(1 - \alpha)$ original confidence interval
                          ($\that_{r} \in [\that_{o} \pm z_{\alpha/2}\sigma_{o}]$).
                          Sometimes, also defined as the original effect estimate is
                          contained in the $(1 - \alpha)$ replication confidence interval
                          ($\that_{o} \in [\that_{r} \pm z_{\alpha/2}\sigma_{r}]$).
                          Usually $\alpha = 5\%$. \\
                   & \\
    Prediction interval ($Q$-test) & The replication effect estimate is contained with its
                                     $(1 - \alpha)$ prediction interval
                                     ($\that_{r} \in [\that_{o} \pm z_{\alpha/2}\surd(\sigma_{o}^{2}
                                     + \sigma_{r}^{2})]$). Usually $\alpha = 5\%$.
                                     This criterion is equivalent to $p_{Q} \geq \alpha$ where
                                     $p_{Q}$ is the $p$-value from the meta-analytic $Q$-test
                                     $p_{Q} = 2\{1 - \Phi(\surd Q)\}$ with $Q$-statistic
                                     $Q = \sum_{i \in \{o, r\}} (\that_{i} - \that_{m})^{2}/\sigma_{i}^{2} = (\that_{o}- \that_{r})^{2}/(\sigma^{2}_{o} + \sigma^{2}_{r})$.\\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not required to show the same direction as this is taken into account by the
$p$-values. Some replication projects \citep{Opensc2015, Errington2021} also
defined replication success via simultaneous non-significance of original and
replication study ($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this
definition ``replication success'' can almost always be achieved by conducting
original and replication study with very few samples so that the $p$-values are
large. The approach is also logically questionable as it could be seen as an
instance of the ``absence of evidence fallacy'' \citep{Altman1995} meaning the
failure to find evidence against the null hypothesis is erroneously interpreted
as evidence for the null hypothesis. Meta-analytic extensions of the
significance approach define replication success by significance of a combined
effect estimate.\footnote{The meta-analytic approach can also be given a
  Bayesian interpretation: A flat prior distribution for the underlying effect
  size $\theta$ is updated by the data from original and replication study.
  Replication success defined by a meta-analytic $p$-value being smaller than
  $\alpha$ is then equivalent to replication success via a Bayesian posterior
  probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
  respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$
  depending on the direction of the combined estimate.} Typically, the
assumption of a common underlying effect size is seen as reasonable so that
fixed-effect meta-analysis is used for pooling. Random-effects meta-analysis has
primarily been used if more than one replication study of the same original
study are conducted since replicators are often interested in understanding
between-replication heterogeneity \citep[\eg{} in][]{Klein2018}.

@@

\subsubsection{Example ``Reproducibility Project: Cancer Biology''}
I will now illustrate the assessment of replicability on data from the
Reproducibility Project: Cancer Biology \citep{Errington2021}. This large-scale
project attempted to replicate 53 landmark studies from the field of cancer
biology \citep[and would perhaps therefore better be named ``Replication
Project'', see][]{Goodman2016}. In the end, only 23 of the 53 studies could be
repeated due to various difficulties, such as, missing information from the
original studies or problems in conducting the experiments.
\citet{Errington2021} report that these experiments led to data on 158
quantitative effects. However, from the data which they provide only 132
quantitative effects are provided with original and replication standardized
mean difference effect estimates along with standard errors, and only they will
be used in the subsequent analyses.

>>>>

\hfill
\begin{tabular}{p{0.6\textwidth}}
  ``\emph{Confirmation comes from repetition}. Any attempt to avoid this statement leads at least to failure and more probably to destruction.'' \\
  \hfill J.W. \citet[p. 84]{Tukey1969}
\end{tabular}

@@

% How can we know if a finding from a scientific study is really true? For
% example, how can we know if the protective effect of a vaccine found in a study
% is real? The answer to such questions is of high importance to scientists,
% decision makers, and the general population. However, because any study result
% comes with uncertainty we can never know for sure.
How can we know whether we can trust a finding from a scientific study? For
example, how can we know whether the protective effect of a vaccine found in a
study is a genuine effect or just a spurious finding? The answer to such
questions is of high importance to scientists, decision makers, and the general
population. However, because every study result is subject to uncertainty, we
can never know for sure.

One way to come a little closer to the truth is to repeat the original study
with new subjects. Such a \emph{replication study} may yield similar results
which would make us more confident about the original finding, or may yield
conflicting results which would lower our confidence. Replication studies are
thus an essential part of the scientific process as they provide a means for
substantiating genuine research findings and refuting research findings which
occurred merely by chance. For this reason, ``successful'' replication is often
a requirement, for instance, for acceptance of newly proposed scientific
theories (\eg{} a new physical model) or the implementation of policies based on
scientific knowledge (\eg{} market approval of a drug). The results from
replication studies may thus have real world consequences, such as deciding
whether we receive the recommendation to get vaccinated (a consequence that
anyone who experienced the COVID-19 pandemic is very aware of).

Despite their tremendous importance, the traditional academic system has made it
unattractive for researchers to conduct replication studies; the currencies of
science -- publications, citations, and grant money -- are typically easier to
acquire by conducting novel research. This is because until recently, most
scientific journals hardly published any replication studies as the emphasis was
on ``innovation'' rather than on ``repetition'' \citep{Makel2012, Martin2017,
  Coiera2021}. To build a successful career, researchers thus often had no other
choice than to concentrate their efforts on producing novel and eye-catching
research results \citep{Binswanger2014, Moher2018}.

@@

Most of these large-scale replication projects confirmed what many researchers
had feared; carefully conducted replication studies often show less impressive
results than their original counterparts, and the replicability of research
findings is surprisingly low (how precisely replicability is defined will soon
be discussed in more detail). This realization led many to declare science as
being in a ``replication crisis''. Debates arose about whether or not the crisis
really existed and who or what was to blame \citep{Gilbert2016, Amrhein2019}.
Even the popular press became interested \citep[\eg{}][]{Carey2015, Kovic2016,
  Achenbach2017, Devlin2018}, so that also in the eyes of the public the
credibility of science became seriously threatened.

\begin{figure}[!htb]
<< "google-scholar-graph", fig.height = 3.5 >>=
## ## Data gathered on October 18 2022
## searchTerm <- "Replication Study"
## searchDate <- as.Date(x = "2022-10-18", format = "%Y-%m-%d", tz = "GMT")
## searchDateFormat <- format(searchDate, format = "%B %d %Y")
## resultsAnytime <- 114000
## scholarDF <- data.frame(year = seq(2022, 1990, -1),
##                         results = c(4530, 11400, 13600, 12200, 11300, 10700,
##                                     9880, 9220, 8660, 8200, 7500, 6700, 6040,
##                                     5370, 4680, 3890, 3380, 3070, 2790, 2520,
##                                     2310, 2190, 2120, 2020, 1920, 1840, 1730,
##                                     1490, 1270, 1230, 1140, 1070, 970),
##                         study = c(347000, 459000, 1710000, 938000, 2360000,
##                                   675000, 3000000, 688000, 3150000, 512000,
##                                   3630000, 451000, 3650000, 396000, 3680000,
##                                   389000, 3890000, 375000, 3950000, 336000,
##                                   4160000, 364000, 4240000, 385000, 3470000,
##                                   377000, 3120000, 335000, 2620000, 376000,
##                                   2440000, 317000, 2220000))
## ## plot data
## ggplot(data = filter(scholarDF, year < 2022),
##        aes(x = year, y = results)) +
##     geom_segment(aes(xend = year, yend = 0), alpha = 0.2, size = 0.4) +
##     geom_point() +
##     ## scale_y_continuous(breaks = seq(0, 15000, 2500)) +
##     scale_x_continuous(breaks = seq(1990, 2020, 5)) +
##     ## scale_y_log10() +
##     labs(x = "Year", y = paste0('Number of results \nfor "', searchTerm, '"')) +
##     theme_bw() +
##     theme(panel.grid.minor = element_blank(),
##           panel.grid.major.x = element_blank()## ,
##           ## panel.grid.major.y = element_line(colour = "grey", linetype = "dashed")
##           )

WebofScienceDF <- data.frame(year = seq(2022, 1990, -1),
                             results = c(139, 144, 121, 137, 111, 80,
                                         89, 63, 68, 92, 63, 60, 39,
                                         39, 41, 30, 21, 20, 19, 25,
                                         15, 9, 13, 20, 19, 16, 14,
                                         8, 17, 13, 21, 5, 12),
                             study = c(235996, 272173, 245994, 230009, 207625,
                                       195012, 174800, 164635, 158745,
                                       184473, 138785, 128779, 116890, 111229,
                                       102231, 95143, 86691, 83132, 94432,
                                       79807, 70388, 53983, 54893, 53480,
                                       51744, 51898, 50070, 46826, 44591,
                                       43097, 42816, 43277, 43277)) %>%
    mutate(proportion = results/study,
           se = sqrt(proportion*(1 - proportion)/study))

## plot data
plotA <- ggplot(data = WebofScienceDF, aes(x = year, y = results)) +
    geom_segment(aes(xend = year, yend = 0), alpha = 0.2, linewidth = 0.5) +
    geom_point(size = 0.7) +
    scale_x_continuous(breaks = seq(1990, 2020, 5)) +
    lims(y = c(0, 160)) +
    labs(x = "Year", y = 'Number of results for "replication study"') +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
plotB <- ggplot(data = WebofScienceDF, aes(x = year, y = proportion)) +
    geom_segment(aes(y = proportion - 2*se, yend = proportion + 2*se, xend = year),
                 size = 0.5, alpha = 0.4) +
    geom_point(size = 0.7) +
    scale_x_continuous(breaks = seq(1990, 2020, 5), name = "Year") +
    scale_y_continuous(labels = scales::percent,
                       name = bquote("Proportion (" %+-% "2SE)")) +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())

ggarrange(plotA, plotB, ncol = 2)
@
\caption{Yearly number of results for search term ``replication study'' on Web
  of Science (left), and number of results normalized by the number of results
  for the search term ``study'' (right). The search was conducted on 2 December
  2022.}
\label{fig:scholar}
\end{figure}

In the midst of this crisis, % , particularly from the hard-hit
% field of psychology,
various reforms were implemented to prevent an escalation of the situation % and
% save the reputation of science
\citep[for an overview see \eg{}][]{Munafo2017}. For instance, many journals
adopted the ``registered report'' format \citep{Chambers2021} in which a study
proposal is peer reviewed \emph{before} the study is conducted, and which, if
accepted, gives provisional publication acceptance regardless of the study
outcome. Similarly, digital infrastructure platforms, such as Zenodo
(\url{https://zenodo.org}) or the Open Science Framework (\url{https://osf.io}),
were created to facilitate preregistration, preprints, code and data sharing,
all of which have substantially increased over the last decade
\citep{Kidwell2016, Nosek2018, Rawlinson2019}. The practice of conducting
replication studies has also gained popularity. Several academic journals and
funders are now explicitly promoting replication research \citep{NWO2016,
  NSL2018, Nature2022}. Figure~\ref{fig:scholar} (left) illustrates this trend
by the yearly number of results for the search term ``replication study''
obtained from the search engine Web of Science. The numbers have been rapidly
growing, especially since the mid-2000s. This trend could possibly be explained
by the fact that more research results are published each year but also when the
numbers are normalized by the number of results for the search term ``study'',
the increasing trend is still visible.

@@

\begin{table}[!htb]
  \centering
  \caption{Study-level summary statistics for an original and a replication
    study. The cumulative distribution function of the standard normal
    distribution is denoted by $\Phi(\cdot)$, and the $1 - \alpha$ quantile of
    the standard normal distribution is denoted by
    $\Phi^{-1}(1 - \alpha) = \z{\alpha}$. Confidence intervals and
    \textit{p}-values are based on the assumption that the effect estimate
    $\that$ is normally distributed around the unknown effect size $\theta$ with
    (known) variance equal to the squared standard error $\sigma^{2}$.}
  \label{tab:meta}
  \begin{tabular}{lcc}
    \toprule
    & Original study & Replication study \\
    \midrule
    effect estimate & $\that_o$ & $\that_{r}$ \\
    standard error & $\sigma_{o}$ & $\sigma_{r}$ \\
    $(1 - \alpha)$ confidence interval &
    $[\that_{o} \pm \z{\alpha/2} \sigma_{o}]$ &
    $[\that_{r} \pm \z{\alpha/2} \sigma_{r}]$ \\
    % $[\that_{o} - \z{\alpha/2} \sigma_{o}, \that_{o} + \z{\alpha/2} \sigma_{o}]$ &
    % $[\that_{r} - \z{\alpha/2} \sigma_{r}, \that_{r} + \z{\alpha/2} \sigma_{r}]$ \\
    $z$-value & $z_{o} = \that_{o}/\sigma_{o}$ & $z_{r} = \that_{r}/\sigma_{r}$ \\
    $p$-value (two-sided) & $p_{o} = 2\{1 - \Phi(|z_{o}|)\}$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    % One-sided $p$-value & $p_{o} = 1 - \Phi(|z_{o}|)$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    \bottomrule
  \end{tabular}
\end{table}

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in research articles, and if one of them is
missing they typically can be calculated from the others. Using summary
statistics is also often the only possible way for conducting the replication
success analysis as the raw data from the original study may not be available to
the researchers conducting the replication study. The most important statistic
is the effect estimate $\that$. It provides an estimate of the underlying effect
size $\theta$ which quantifies the true effect or association of an
intervention/exposure with an outcome variable. Typical effect sizes are mean
differences and correlations (for continuous outcomes), odds ratios (for binary
outcomes), or hazard ratios (for time to event outcomes). Depending on the type
of effect size, a transformation might be required so that the assumption of
approximately normally distributed effect estimates around the unknown effect
size $\theta$ (for large enough sample sizes) is justified. This could be, for
instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard error $\sigma$ represents the statistical uncertainty of the
estimate.\footnote{Typically, the available standard error $\sigma$ is only an
  estimate of the true standard error. However, in most parts of this thesis it
  will be assumed that $\sigma$ corresponds to the true standard error. This is
  the same assumption as in ordinary fixed-effects/random-effects meta-analysis,
  and reasonable if the sample size of the studies is not too small.} Under the
assumption of (asymptotic) normality, confidence intervals for $\theta$ and
$p$-values for testing the null hypothesis $H_{0} \colon \theta = 0$ can be
computed as shown in Table~\ref{tab:meta}.

\begin{table}[!htb]
  \centering
  \caption{Statistical criteria for assessing replication success which have
    been used in practice \citep{Opensc2015, Camerer2016, Camerer2018, Cova2018,
      Errington2021}.}
  \label{tab:rsdef}
  \begin{tabular}{p{0.21\textwidth}p{0.7\textwidth}}
    \toprule
    Criterion type & Description \\
    \midrule
    Significance & The original and replication $p$-values are smaller than
                   a threshold $\alpha$ and their effect estimates show
                   the same direction
                   (\mbox{$p_{o} < \alpha$}, \mbox{$p_{r} < \alpha$}, and
                   \mbox{$\sign(\that_{o}) = \sign(\that_{r})$}). Usually $\alpha = 5\%$.
    \\
                   & \\
    Meta-analytic significance & The meta-analytic $p$-value is smaller than a threshold
                                 $\alpha$ (\mbox{$p_{m} = 2\{1 - \Phi(|\that_{m}|/\sigma_{m})\} < \alpha$})
                                 where \mbox{$\that_{m} =
                                 (\that_{o}/\sigma_{o}^{2} + \that_{r}/\sigma^{2}_{r})\sigma^{2}_{m}$}
                                 is the  pooled effect estimate with standard error
                                 \mbox{$\sigma_{m} = (1/\sigma^{2}_{o} + 1/\sigma^{2}_{r})^{-1/2}$}.
                                 Usually $\alpha = 5\%$. \\
                   & \\
    Relative effect size & The effect estimate of the replication study goes in the same
                           direction as the original one and is at least $d_{\text{min}}$ times as large (\mbox{$d = \that_{r}/\that_{o} \geq d_{\text{min}}$}).
                           Usually $d_{\text{min}} = 1$. \\
                   & \\
    Confidence interval & The replication effect estimate is contained in the
                          $(1 - \alpha)$ original confidence interval
                          ($\that_{r} \in [\that_{o} \pm z_{\alpha/2}\sigma_{o}]$).
                          Sometimes, also defined as the original effect estimate is
                          contained in the $(1 - \alpha)$ replication confidence interval
                          ($\that_{o} \in [\that_{r} \pm z_{\alpha/2}\sigma_{r}]$).
                          Usually $\alpha = 5\%$. \\
                   & \\
    Prediction interval ($Q$-test) & The replication effect estimate is contained in its
                                     $(1 - \alpha)$ prediction interval
                                     ($\that_{r} \in [\that_{o} \pm z_{\alpha/2}\surd(\sigma_{o}^{2}
                                     + \sigma_{r}^{2})]$). Usually $\alpha = 5\%$.
                                     This criterion is equivalent to $p_{Q} \geq \alpha$ where
                                     $p_{Q}$ is the $p$-value from the meta-analytic $Q$-test
                                     $p_{Q} = 2\{1 - \Phi(\surd Q)\}$ with $Q$-statistic
                                     $Q = \sum_{i \in \{o, r\}} (\that_{i} - \that_{m})^{2}/\sigma_{i}^{2} = (\that_{o}- \that_{r})^{2}/(\sigma^{2}_{o} + \sigma^{2}_{r})$.\\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not required to show the same direction as this is taken into account by the
one-sided $p$-values. Some replication projects \citep{Opensc2015,
  Errington2021} also defined replication success via simultaneous
non-significance of original and replication study ($p_{o} > \alpha$ and
$p_{r} > \alpha$). However, with this definition ``replication success'' can
almost always be achieved by conducting original and replication study with very
few samples so that the $p$-values are large. The approach is also logically
questionable as it could be seen as an instance of the ``absence of evidence
fallacy'' \citep{Altman1995} meaning the failure to find evidence against the
null hypothesis is erroneously interpreted as evidence for the null hypothesis.
Meta-analytic extensions of the significance approach define replication success
by significance of a combined effect estimate.\footnote{The meta-analytic
  approach can also be given a Bayesian interpretation: A flat prior
  distribution for the underlying effect size $\theta$ is updated by the data
  from original and replication study. Replication success defined by a
  meta-analytic $p$-value being smaller than $\alpha$ is then equivalent to
  replication success via a Bayesian posterior probability
  $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$, respectively,
  $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$, depending on the
  direction of the combined estimate.} Typically, the assumption of a common
underlying effect size is seen as reasonable so that fixed-effect meta-analysis
is used for pooling. Random-effects meta-analysis has primarily been used if
more than one replication study of the same original study are conducted since
replicators are often interested in understanding between-replication
heterogeneity \citep[\eg{} in][]{Klein2018}.

@@

\subsubsection{Example ``Reproducibility Project: Cancer Biology''}
I will now illustrate the assessment of replication success on data from the
``Reproducibility Project: Cancer Biology'' \citep{Errington2021}. This
large-scale project attempted to replicate 53 landmark studies from the field of
cancer biology \citep[and would perhaps therefore better be named ``Replication
Project'', see][]{Goodman2016}. In the end, only 23 of the 53 studies could be
repeated due to various difficulties, such as, missing information from the
original studies or problems in conducting the experiments.
\citet{Errington2021} report that these experiments led to data on 158
quantitative effects. However, from the data which they provide only 132
quantitative effects are provided with original and replication standardized
mean difference effect estimates along with standard errors, and only they will
be used in the subsequent analyses.

\begin{figure}[!htb]
<< "RPCB-graph", fig.height = 3.5 >>=

## function to format CI
ciFormat. <- function(lower, upper) {
    paste0("[", format(signif(lower, 2), nsmall = 2),
           ", ", format(signif(upper, 2), nsmall = 2), "]")
}
ciFormat <- Vectorize(FUN = ciFormat.)

## function to format p-value
pFormat. <- function(p, digits = 2, dollar = FALSE, phantom = FALSE) {
    if (is.na(p)) {
        out <- NA
    } else {
        if (p > 0.0001) {
            out <- as.character(signif(p, digits = digits))
            if (phantom) out <- paste0("\\phantom{<~}", out)
        } else {
            out <- "< 0.0001"
        }
        if (dollar) {
            out <- paste0("$", out, "$")
        }
    }
    return(out)
}
pFormat <- Vectorize(FUN = pFormat.)

## function to format (relative) effect estimate
dFormat. <- function(x, digits = 2, phantom = FALSE, dollar = FALSE) {
    if (is.na(x)) {
        out <- NA
    } else {
        out <- as.character(signif(x, digits = digits))
        if (phantom && (x >= 0)) {
           out <- paste0("\\phantom{-}", out)
        }
        if (dollar) out <- paste0("$", out, "$")
    }
    return(out)
}
dFormat <- Vectorize(FUN = dFormat.)

## function to color code numerical value in LaTeX table based on threshold
tableFormat. <- function(x, outFun = pFormat, threshold = 0.05, col1 = "nicegreen",
                         col2 = "red") {
    if (is.na(x)) {
        out <- NA
    } else {
        if (x < threshold) col <- col1
        else col <- col2
        out <- paste0("\\textcolor{", col, "}{",
                      outFun(x, dollar = TRUE, phantom = TRUE), "}")
    }
    return(out)
}
tableFormat <- Vectorize(FUN = tableFormat., vectorize.args = "x")
